---
title: "Lab1_JunLi"
subtitle: "Computational Statistics -- 732A90"
author: "Jun Li"
date: '2020-02-12'
output: pdf_document
---

```{r,eval=TRUE,echo=FALSE,warning=FALSE,message=FALSE}
RNGversion(min(as.character(getRversion()),"3.6.2")) ## with your R-version
#set.seed(12345, kind = "Mersenne-Twister", normal.kind = "Inversion")
```



# Question 1: Cluster Sampling

It shows that most of the selected cities have large populations, which is consistent with the positive relationship between population and sampling probability.

The two histograms show an approximately similar distribution, where the cities with larger population are likely to be included in sampling. Therefore the sampling function meets the requirement.



```{r,eval=TRUE,echo=FALSE,warning=FALSE,results=FALSE}
# Part 1
da1<-read.csv2("population.csv")


# Part 2
#myrunif<-function(a,c,m,x0,n)
#   {re<-rep(0,n)
#    for(i in 1:n){
#      if(i==1) re[i]<-(a*x0+c)%%m
#      if(i>1) re[i]<-(a*re[i-1]+c)%%m}
  
#    return (re/m)}

mycity<-function(den)
  {n<-length(den)
   cdf<-rep(0,n)
   for(i in 1:n) cdf[i]<-sum(den[1:i])
   ran<-runif(1)                # allowed to use runif() ???
   return (which(cdf>ran)[1])   # the index of selected city
   }

# Part 3
cities<-data.frame(Municipality=rep("City",20),Population=rep(0,20),stringsAsFactors=FALSE)
da<-da1;den<-da[,2]/sum(da[,2])

for(i in 1:20)
  {ind<-mycity(den)
   cities[i,1]<-as.character(da[ind,1])
   cities[i,2]<-da[ind,2]
   da<-da[-ind]
   den<-da[,2]/sum(da[,2])}

# Part 4
print("Here come the selected cities: ")
cities

# Part 5
hist(da1[,2],main="All cities",labels=da1[,1],breaks=100)
hist(cities[,2],main="Selected cities",breaks=100)
```

# Question 2: Different distributions
## Part 1
The goal is to get inverse of CDF of original random variable with uniform random generated numbers within (0,1) as input:
$$F(x)=\int_{-\infty}^{x}\frac{1}{2}e^{-|x|}$$

if x<0, $$F(x)=\frac{1}{2}e^{x}$$
$$F^{-1}(U)=ln2U  (0<U<0.5)$$
if x>=0,$$F(x)=1-\frac{1}{2}e^{-x}$$ 
$$F^{-1}(U)=-ln2(1-U)  (0.5=<U<1)$$
The simulations from the formulars above seem to work well.

## Part 2
The Acceptance/rejection method tries to find a minimal c which meets the following requirement:
$$c*f_Y(x)>=f_X(x)$$
$$c>=\frac{f_X(x)}{f_Y(x)}=\frac{\frac{1}{\sqrt{2\pi}}e^{-\frac{x^2}{2}}}{\frac{1}{2}e^{-|x|}}=\sqrt{\frac{2}{\pi}}e^{-\frac{x^2}{2}+|x|}=\sqrt{\frac{2}{\pi}}e^{-\frac{x^2}{2}-x}(x<0)<=\sqrt{\frac{2}{\pi}}e^{0.5}=1.315489$$
$$=\sqrt{\frac{2}{\pi}}e^{-\frac{x^2}{2}+x}(x>=0)=\sqrt{\frac{2}{\pi}}e^{0.5}=1.315489$$
The histograms of simulated numbers from acceptance/rejection method and normal random numbers show the significantly consistent distributions.

```{r,eval=TRUE,echo=FALSE,warning=FALSE,message=FALSE}
# Part 1
mysim<-function()
  {ran<-runif(1)
   if(ran<0.5) sim<-log(2*ran) else
               sim<--log(2*(1-ran))
   return (sim)}

sim<-rep(0,10000)
for(i in 1:10000) sim[i]<-mysim()
hist(sim,main="Simulations",breaks=100)

# Part 2
myacrec<-function(c)
  {x<-NULL
   reject<-0
   while(is.null(x))
   {Y<-mysim()
    U<-runif(1)
    if(U<=dnorm(Y,0,1)/(c*(0.5*exp(-abs(Y))))) x<-Y else
      reject<-reject+1}
  return (c(x,reject))}

sim1<-rep(0,2000)
sumreject<-0
for(i in 1:2000)  {re<-myacrec(1.315489);sim1[i]<-re[1];sumreject<-sumreject+re[2]}
hist(sim1,main="Simulated N(0,1) with acceptance/rejection method",breaks=100)
print(paste("The average rejection rate R is: ",sumreject/(2000+sumreject)," which is approximately equal with ",1-1/1.315489,sep=""))
 
xnorm<-rnorm(2000,0,1)
hist(xnorm,main="Normal random numbers",breaks=100)


```


# Code Appendix

```{r code = readLines(knitr::purl("C:/Users/A550240/Desktop/LIU/2.1_CompuStatistics/Labs/Lab3_JunLi.Rmd", documentation = 1)), echo = T, eval = F}
```