---
title: "Lab4_JunLi"
subtitle: "Computational Statistics -- 732A90"
author: "Jun Li"
date: '2020-02-17'
output: pdf_document
---

```{r,eval=TRUE,echo=FALSE,warning=FALSE,message=FALSE}
RNGversion(min(as.character(getRversion()),"3.6.2")) ## with your R-version
#set.seed(12345, kind = "Mersenne-Twister", normal.kind = "Inversion")
```



# Question 1: Hypothesis testing
## Part 1.1/2
The scatter plot presents a lookingly random behaviour of lottery. However, the fitted values from Loess() shows a slightly negative relationship between X and expected Y.

```{r,eval=TRUE,echo=FALSE,warning=FALSE,results=FALSE}
# Part 1
da1<-read.csv2("lottery.csv")
Y<-da1$Draft_No;X<-da1$Day_of_year
# Part 2
fit<-loess(Y~X)
Y1<-fit$fitted
plot(X,Y)
lines(X,Y1,col="red",add=TRUE)
```



## Part 1.3
P-value from the hypothesis testing is 0.0035<1%, which means that "H0:T is significantly larger than 0" is rejected. Therefore, there is no trend in lottery and it is random.

```{r,eval=TRUE,echo=FALSE,warning=FALSE,results=FALSE}
# Part 3
set.seed(12345, kind = "Mersenne-Twister", normal.kind = "Inversion")
B<-2000;   ## input
n<-dim(da1)[1]
t<-rep(NULL,B)
for(i in 1:B){
  samx<-sample(1:n,n,replace=TRUE)
  samy<-da1[samx,]$Draft_No
  samda<-cbind(samx,samy);#names(samda)=c("samx","samy")
  samyfit<-loess(samy~samx)$fitted
  t[i]<-(max(samyfit)-min(samyfit))/(samx[which.max(samyfit)]-samx[which.min(samyfit)])}
hist(t,breaks=100,main="Test 1.3")
print(paste("P value is: ",mean(t>0),sep=""))
```



## Part 1.4
P-value of two-sided test is 0.16>5%, which means that H0 can not be rejected. Therefore, permutation test shows that the lottery is random.

```{r,eval=TRUE,echo=FALSE,warning=FALSE,results=FALSE}
# Part 4
myperm<-function(B,data){
set.seed(12345, kind = "Mersenne-Twister", normal.kind = "Inversion")
n<-dim(data)[1]
t<-rep(NULL,B)

for(i in 1:B){
  permx<-sample(1:n,n)
  y<-data$Draft_No
  permda<-cbind(permx,y);#names(samda)=c("samx","samy")
  permyfit<-loess(y~permx)$fitted
  t[i]<-(max(permyfit)-min(permyfit))/(permx[which.max(permyfit)]-permx[which.min(permyfit)])}

fit0<-loess(Draft_No~Day_of_year,data=data)$fitted
t0<-(max(fit0)-min(fit0))/(data$Day_of_year[which.max(fit0)]-data$Day_of_year[which.min(fit0)])
p<-mean(abs(t)>abs(t0))  ## two-sided

return (list(pvalue=p,originalt=t0,samplet=t))}

re<-myperm(2000,da1)
p<-re$pvalue
t0<-re$originalt
t<-re$samplet
hist(t,breaks=100,main="Test 1.4")
abline(v=t0,col="blue")
print(paste("P value is: ",p,sep=""))
```


## Part 1.5
With 100 loops for each alpha, the series of power index is presented in the graph. Each alpha has a power of 1, which confirms a high quality of test statistic.

```{r,eval=TRUE,echo=FALSE,warning=FALSE,results=FALSE}
# Part 5
#set.seed(12345, kind = "Mersenne-Twister", normal.kind = "Inversion")

generate<-function(alpha){   ## generate dataset
  X<-da1$Day_of_year
  beta<-rnorm(366,183,10)
  Y<-pmax(0,pmin(alpha*X+beta,366))
  return (as.data.frame(cbind(Day_of_year=X,Draft_No=Y)))}

alpha<-seq(0.1,10,0.1)
n<-length(alpha)
power<-rep(NULL,n)
for(i in 1:n){
  err2<-0
  for(j in 1:100){
    da<-generate(alpha[i])
    p<-myperm(200,da)$pvalue
    if(p>=0.05) err2<-err2+1}
  power[i]<-1-err2/100}

plot(alpha,power,main="Power trend")
```



# Question 2: Bootstrap, jackknife and confidence intervals
## Part 1
The price histogram suggests that it probably follows the gamma distribution.

```{r,eval=TRUE,echo=FALSE,warning=FALSE,message=FALSE}
# Part 1
da2<-read.csv2("prices1.csv")
hist(da2$Price,breaks=100,prob=TRUE)
lines(density(da2$Price),col="blue")
print(paste("The mean price is: ",mean(da2$Price),sep=""))
```

## Part 2
The estimation of mean price distribution is as follows:

```{r,eval=TRUE,echo=FALSE,warning=FALSE,message=FALSE}
# Part 2
library("boot")

stat1<-function(data,vn){
  data<-as.data.frame(data[vn,])
  return (mean(data$Price))}

res<-boot(da2,stat1,R=1000)
boot.ci(res)  ## confidence intervals with different methods
plot(res)

print(paste("The estimated variance of statistic with bootstrap is: ",sum((res$t-mean(res$t))^2)/(1000-1),sep=""))
print(paste("The bias-corrected mean of statistic with bootstrap is: ",2*res$t0-mean(res$t),sep=""))
```

## Part 3
The estimated variance of mean price with jackknife is larger than bootstrap method



```{r,eval=TRUE,echo=FALSE,warning=FALSE,message=FALSE}
# Part 3
t0<-mean(da2$Price)
n<-nrow(da2)
t<-rep(NULL,n)
for(i in 1:n) t[i]<-mean(da2$Price[-i])

print(paste("The jackknife mean price is: ",sum((n*t0-(n-1)*t-mean(t))^2)/(n*(n-1)),sep=""))
```



## Part 4
Confidence intervals with different methods are ploted in the graph as below, where the Normal method has largest length and smallest center location, while Percentile and BCa methods share a similar interval.


```{r,eval=TRUE,echo=FALSE,warning=FALSE,message=FALSE}
# Part 4
w<-10
ci<-boot.ci(res)

cat(paste(sprintf("%*s",w,""),sprintf("%*s",w,"lower")),sprintf("%*s",w,"upper"),sprintf("%*s",w,"center"),sprintf("%*s",w,"length"))
cat("\n")
cat(paste(sprintf("%*s",w,"Percentile"),sprintf("%*.7s",w,ci$percent[4]),sprintf("%*.7s",w,ci$percent[5]),sprintf("%*.7s",w,(ci$percent[4]+ci$percent[5])/2),sprintf("%*.7s",w,ci$percent[5]-ci$percent[4])))
cat("\n")
cat(paste(sprintf("%*s",w,"BCa"),sprintf("%*.7s",w,ci$bca[4]),sprintf("%*.7s",w,ci$bca[5]),sprintf("%*.7s",w,(ci$bca[4]+ci$percent[5])/2),sprintf("%*.7s",w,ci$bca[5]-ci$bca[4])))
cat("\n")
cat(paste(sprintf("%*s",w,"Normal"),sprintf("%*.7s",w,ci$normal[2]),sprintf("%*.7s",w,ci$normal[3]),sprintf("%*.7s",w,(ci$normal[2]+ci$normal[3])/2),sprintf("%*.7s",w,ci$normal[3]-ci$normal[2])))
```


# Code Appendix

```{r code = readLines(knitr::purl("C:/Users/A550240/Desktop/LIU/2.1_CompuStatistics/Labs/Lab5_JunLi.Rmd", documentation = 1)), echo = T, eval = F}
```