---
title: "Lab6_JunLi"
subtitle: "Computational Statistics -- 732A90"
author: "Jun Li"
date: '2020-03-06'
output: pdf_document
---

```{r,eval=TRUE,echo=FALSE,warning=FALSE,message=FALSE}
RNGversion(min(as.character(getRversion()),"3.6.2")) ## with your R-version
#set.seed(12345, kind = "Mersenne-Twister", normal.kind = "Inversion")
```



# Question 1: Genetic algorithm
## Part 1.1-1.4
code for 1.1~1.4 refer to the appendix.

```{r,eval=TRUE,echo=FALSE,warning=FALSE,results=FALSE}
# Part 1.1
f<-function(x) return (x^2/exp(x)-2*exp(-9*sin(x)/(x^2+x+1)))

# Part 1.2
crossover<-function(x,y) return ((x+y)/2)

# Part 1.3
mutate<-function(x) return (floor(x^2%%30)) 

# Part 1.4
myfunc<-function(maxiter,mutprob){
  curve(f,0,30,xlab="x",ylab="Base function f",col="black",main=paste(maxiter,"/",mutprob,sep=""))
  X<-seq(0,30,5)  # initial population
  Values<-f(X)
  points(X,Values,typ='p',col='blue',lwd=15,add=TRUE)
  max<--Inf;maxind<-NULL
  
  for(i in 1:maxiter){ # reproduction
    parents<-sample(X,2)
    victim<-order(Values)[1]   #min.which(Values)
    
    prekid<-crossover(parents[1],parents[2])
    kid<-mutate(prekid)*mutprob+(1-mutprob)*prekid
    X[victim]<-kid
    Values[victim]<-f(kid)
    
    tempmax<-max(Values)
    if(tempmax>max) {max<-tempmax;maxind<-X[which.max(Values)]}
    }
    
  points(maxind,max,typ='p',col='red',lwd=15,add=TRUE)
  legend('topright',legend=c("Objective func","Initial popu","Maxi"),fill=c("black","blue","red"))
  return (max)}


myfunc(10,0.1)

```



## Part 1.5
The collection of plots show that geneti algorithm works well with maximum optimization. Particularly, the higher mutation probability truely helps exit a local optimal.

```{r,eval=TRUE,echo=FALSE,warning=FALSE,results=FALSE}
# Part 1.5
maxiter<-c(10,100)
mutprob<-c(0.1,0.5,0.9)
par(mfrow=c(3,2),mar=c(2,2,2,2))
for(i in 1:2)
  for(j in 1:3)
    myfunc(maxiter[i],mutprob[j])

```





# Question 2: EM algorithm
## Part 1
The plot shows a high correlation between Y and Z versus X, and the variance of Y and Z is decreasing through X.

```{r,eval=TRUE,echo=FALSE,warning=FALSE,message=FALSE}
# Part 1
da<-read.csv("physical1.csv")
X<-da$X;Y<-da$Y;Z<-da$Z
plot(X,Y,typ="l",col="blue",ylim=c(0,30));lines(X,Z,typ="l",col="red")
legend('topright',legend=c("Y/X","Z/X"),fill=c("blue","red"))
```

## Part 2
Deduction refer to appendix of notes.

## Part 3


```{r,eval=TRUE,echo=FALSE,warning=FALSE,message=FALSE}
# Part 3
n<-length(X);
nm<-sum(is.na(Z));no<-n-nm
indm<-which(is.na(Z))

lambda<-100
repeat{
  nylambda<-(2*sum(X*Y)+sum(X[-indm]*Z[-indm])+2*nm*lambda[length(lambda)])/(4*n)
  lambda<-c(lambda,nylambda)
  if(abs(lambda[length(lambda)]-lambda[length(lambda)-1])<0.001) break()
}

re<-lambda[length(lambda)]
print(paste("The optimal lambda is: ",re,sep=""))
print(paste("The number iterations is: ",length(lambda)-1,sep=""))


```



## Part 4
The expected Y and Z seem to follow the trend of actual values, which confirmed the effect of EM algorithm.

```{r,eval=TRUE,echo=FALSE,warning=FALSE,message=FALSE}
# Part 4
ey<-re/X
ez<-2*re/X

plot(X,Y,typ="l",col="blue",ylim=c(0,30));lines(X,Z,typ="l",col="red")
lines(X,ey,typ="l",col="green");lines(X,ez,typ="l",col="brown")
legend('topright',legend=c("Y/X","Z/X","E(Y)","E(Z)"),fill=c("blue","red","green","brown"))


```


# Code Appendix

```{r code = readLines(knitr::purl("C:/Users/A550240/Desktop/LIU/2.1_CompuStatistics/Labs/Lab6_JunLi.Rmd", documentation = 1)), echo = T, eval = F}
```